Reanalysis Step 2: Multiple Imputation

### 2025-08-30 reanalysis-2-multiple_imputation.qmd
cci and dcsi consistency

@@

### Person-Level DR Consistency Check

```{r dr-consistency-check}
# Check that person_dr is consistent within patients after imputation
dr_consistency_check <- lapply(1:n_imputations, function(i) {
  imp_data <- imputed_datasets[[i]]

  # Check consistency of person_dr within patients
  dr_check <- imp_data |>
    group_by(e_mrn_deidentified) |>
    summarise(
      n_eyes = n(),
      unique_dr_values = n_distinct(person_dr, na.rm = TRUE),
      is_consistent = unique_dr_values <= 1  # Should be 1 unique value per patient
    )

  n_inconsistent <- sum(!dr_check$is_consistent)
  return(list(
    dataset = i,
    n_patients_checked = nrow(dr_check),
    n_inconsistent_patients = n_inconsistent,
    pct_consistent = round((1 - n_inconsistent/nrow(dr_check)) * 100, 2)
  ))
})

# Display results
dr_consistency_df <- bind_rows(dr_consistency_check)
log_message(paste0("DR consistency check: ", 
                  min(dr_consistency_df$pct_consistent), "% to ",
                  max(dr_consistency_df$pct_consistent), "% consistent across datasets"))
```

**Person-level DR consistency after imputation:**
```{r dr-consistency-table}
kable(dr_consistency_df, 
      col.names = c("Dataset", "Patients", "Inconsistent", "% Consistent"))
```

>>>>

# FINAL STEP: Enforce patient-level variable consistency across all datasets
# person_dr, CCI, and DCSI should be consistent within patients
imputed_datasets <- lapply(imputed_datasets, function(imp_data) {
  # Calculate consistent values per patient (using max for all)
  patient_level_vars <- imp_data |>
    group_by(e_mrn_deidentified) |>
    summarise(
      person_dr_consistent = max(person_dr, na.rm = TRUE),
      CCI_consistent = max(CCI, na.rm = TRUE),
      DCSI_consistent = max(DCSI, na.rm = TRUE),
      .groups = 'drop'
    )

  # Replace with consistent values
  imp_data <- imp_data |>
    left_join(patient_level_vars, by = "e_mrn_deidentified") |>
    mutate(
      person_dr = person_dr_consistent,
      CCI = CCI_consistent,
      DCSI = DCSI_consistent
    ) |>
    select(-person_dr_consistent, -CCI_consistent, -DCSI_consistent)

log_message("Enforced patient-level consistency for person_dr, CCI, and DCSI by taking max value per patient")

@@

### Patient-Level Variable Consistency Check

```{r patient-consistency-check}
# Check that patient-level variables are consistent within patients after imputation
consistency_check <- lapply(1:n_imputations, function(i) {
  imp_data <- imputed_datasets[[i]]

  # Check consistency of person_dr, CCI, and DCSI within patients
  patient_check <- imp_data |>
    group_by(e_mrn_deidentified) |>
    summarise(
      n_eyes = n(),
      unique_dr = n_distinct(person_dr, na.rm = TRUE),
      unique_cci = n_distinct(CCI, na.rm = TRUE),
      unique_dcsi = n_distinct(DCSI, na.rm = TRUE),
      dr_consistent = unique_dr <= 1,
      cci_consistent = unique_cci <= 1,
      dcsi_consistent = unique_dcsi <= 1,
      all_consistent = dr_consistent & cci_consistent & dcsi_consistent
    )

  return(list(
    dataset = i,
    n_patients = nrow(patient_check),
    dr_pct_consistent = round(mean(patient_check$dr_consistent) * 100, 2),
    cci_pct_consistent = round(mean(patient_check$cci_consistent) * 100, 2),
    dcsi_pct_consistent = round(mean(patient_check$dcsi_consistent) * 100, 2),
    all_pct_consistent = round(mean(patient_check$all_consistent) * 100, 2)
  ))
})

# Display results
consistency_df <- bind_rows(consistency_check)
log_message(paste0("Patient-level consistency - person_dr: ", 
                  min(consistency_df$dr_pct_consistent), "%, ",
                  "CCI: ", min(consistency_df$cci_pct_consistent), "%, ",
                  "DCSI: ", min(consistency_df$dcsi_pct_consistent), "%"))
```

**Patient-level variable consistency after imputation:**
```{r consistency-table}
kable(consistency_df, 
      col.names = c("Dataset", "Patients", "DR %", "CCI %", "DCSI %", "All %"),
      caption = "Percentage of patients with consistent values across eyes")
```
